How to compute the fusion product of MV cycles in type A

### 2021-03-31 draft.tex
minor moves

    \begin{remark}
        Base case $m = 1$: by a slight abuse of notation let $\lambda_1, \lambda_2 \in \NN$. Claim: $J=J_{0,\lambda_1} \oplus J_{s,\lambda_2}$ is conjugate to

% $\begin{pNiceArray}{CC|CC}[first-row,last-row=5,first-col,last-col,nullify-dots]
% & C_1 & \Cdots & & C_4 & \\
% L_1 & a_{11} & a_{12} & a_{13} & a_{14} & L_1 \\
% \Vdots & a_{21} & a_{22} & a_{23} & a_{24} & \Vdots \\
% \hline
% & a_{31} & a_{32} & a_{33} & a_{34} & \\
% L_4 & a_{41} & a_{42} & a_{43} & a_{44} & L_4 \\
% & C_1 & \Cdots & & C_4 &
% \end{pNiceArray}$
\begin{comment}
\[
        A = \begin{bNiceArray}{ccc|ccc}[columns-width = auto,first-row,last-col,code-for-first-row = \color{blue}\scriptstyle\rotate,code-for-last-col = \color{blue}\scriptstyle,nullify-dots,xdots/line-style=loosely dotted]
        1 & \Cdots & \lambda_1 & \lambda_1 + 1 & \Cdots & \lambda_1 + \lambda_2 \\
        0 & 1 & 0 & 0 & 0 & 0 & 1\\
        0 & \Ddots & 1 & 0 & 0 & 0 & \Vdots \\
        0 & 0 & 0 & 1 & 0 & 0 & \lambda_1 \\
        \hline
         & & & s & 1 & 0 & \lambda_1 + 1 \\
         & & & 0 & \Ddots & 1 & \Vdots \\
         & & & 0 & 0 & s & \lambda_1 + \lambda_2 
        \end{bNiceArray} 
\]
\end{comment}
% Denote by $R_i$ the $i$th row of $J$ 
%Let $g = E_{\lambda_2}\cdots E_1$ be the product of elementary matrices $E_i$ which correspond to the elementary row operations 
%$$R_{\lambda_1 + i - 1} \leftarrow R_{\lambda_1 + i-1} + \frac 1s R_{\lambda_1 + i}$$ 
%for $i = 1,\dots,\lambda_2$. Then $gJg^{-1} = A$. For the next part we introduce the notation $A_{\lambda_1+\lambda_2}$ for this matrix $A$.

@@

\acom{The following is a rough note.}

The relevant result from Eisenbud--Saltman seems to be that 
$$X_r = X_{r,0} = \{A:\rk A^i \le r(i) \}$$ 
is the flat limit of 
$$X_{r,\lambda} = \{A : \rk (A-\lambda_i)^j\le r(i,j)\}$$
with $r(i,j)$ determined by $r(i)$.

The case that we are interested in is the case of two eigenvalues. Let me change notation so that $\lambda_i$ are shapes and let me only look at the open loci where the rank conditions are strict.

Let $m_i = \lambda_{i,1}$ ($i = 1,2$), $m = m_1 + m_2$, and define
$$r(j) = N - \#\text{~boxes in first $j$ columns of }\lambda_1 + \lambda_2 = \lambda\vdash N$$
on $j = 1,\dots,m$. Note that $r(j-1) - r(j)$ is the length of the $j$th column of $\lambda$. As in \cite{eisenbud1989rank} we will denote this quantity $a_j$. 
%

% Again too small. The problem is I don't think that it is in general true that some permutation of the list $r(i-1) - r(i)$ will be the list of lengths of columns of $\lambda_1$ followed by the list of lengths of columns of $\lambda_2$ which is what we are looking for. For one thing there's in general more than one way to write a partition as a sum of two partitions. Ok, concretely, a third example. Take $\lambda_1 = (a,b)$ and $\lambda_2 = (c,d)$ then $\lambda = (a + c, b + d)$ and 
% $$r\{0,1,\dots,a + c\} = \{N, N-2, \dots, N-2(b+d), N-2(b+d) - 1,\dots, 0\}$$
% so 
% $$\{a_1,a_2,\dots\} = \{2,2,\dots,2,1,1,\dots,1\}$$
% with $b+d$ 2s and $a+c - b - d$ 1s. Wait a minute, nevermind, of course this is true. We have two partitions $\lambda_1$ and $\lambda_2$ and one way to form the partition $\lambda = \lambda_1 + \lambda_2$ is to break $\lambda_i$ down into columns, order the columns by length, and recombine. Cool. I think it's fine. 
\acom{That's it for the note.}

@@

% If we apply the change of basis $g$ to the ideal $I=I_1 + I_2$ and set $s = 0$ (if possible) do we recover the ideal of $\OO^{\lambda_1 + \lambda_2}$? Let $A^g = g(A_1 \oplus A_2) g^{-1}$...
    \end{remark}

@@

@@

    Now assume that $ (L, s) \in \Gr^{\lambda_1, \lambda_2}_{0,s}$. We know that \acom{Rather than we know, it is by definition?} $ L(0) \in \Gr^{\lambda_1}$ and $L(s) \in \Gr^{\lambda_2} $.  This means that $t $ acting on 
    % $ \CC\xt^m / L(0)$ 
    $\cO^m/L(0)$ has Jordan type $ \lambda_1$ and $ t$ acting on 
    % $ \CC\xt[t-s]^m / L(s) $ 
    $\cO_s^m/L(s)$ has Jordan type $ \lambda_2$.  Now, the map 
    % $ L_0 / L \rightarrow \CC\xt^m / L(0)$ 
    $\CC[t]^m/L\to \cO^m/L(0)$ induces an isomorphism between the $0$-generalized eigenspace of $ t$ and $ \cO^m / L(0)$.   The same thing holds for the $s$-generalized eigenspace of $t $ and $ \cO_s^m/L(s)$. This shows that (i) implies (ii) and the logic can be reversed to see that (ii) implies (i). \acom{is this map $v + L \mapsto v + L(0)$?}

@@

    the defining fact that lattices satisfying Equation~\ref{eq:defGrBDlambda} equivalently satisfy 
    \begin{equation}
        \begin{split}
            t\big|_{\CC\xt^m/L_1}\text{ has Jordan type} \le \lambda_1 \\
            t\big|_{\CC\xt^m/L_2}\text{ has Jordan type} \le \lambda_2 
        \end{split}
    \end{equation}
    where recall 
    $L_i = L\otimes \CC\xt$ ??? 
    % $L_i = L\otimes \CC[(t-p_i)^{-1}]$ 
    and $p_1 = s$ while $p_2 = 0$. \acom{Somehow, restricting to an eigenspace is like inverting/forgetting the action of $t$ by any other generalized eigenvalue? Basic linear algebra? Joel?}
\end{enumerate}
    \end{enumerate}
\end{proof}

@@

    Need: operations of taking quotients and taking generalized eigenspaces (i.e.\ $-\otimes\CC\xt[t-a]$) commute.

    \acom{Does this have to do with the fact that a functor $F: \sch \to \set$ commutes with projective limit if and only if it's representable?}
\end{proof}

>>>>

% TODO: delete commented out remark and subcomment!
% \begin{remark}
\begin{comment}
        Base case $m = 1$: by a slight abuse of notation let $\lambda_1, \lambda_2 \in \NN$. Claim: $J=J_{0,\lambda_1} \oplus J_{s,\lambda_2}$ is conjugate to

% $\begin{pNiceArray}{CC|CC}[first-row,last-row=5,first-col,last-col,nullify-dots]
% & C_1 & \Cdots & & C_4 & \\
% L_1 & a_{11} & a_{12} & a_{13} & a_{14} & L_1 \\
% \Vdots & a_{21} & a_{22} & a_{23} & a_{24} & \Vdots \\
% \hline
% & a_{31} & a_{32} & a_{33} & a_{34} & \\
% L_4 & a_{41} & a_{42} & a_{43} & a_{44} & L_4 \\
% & C_1 & \Cdots & & C_4 &
% \end{pNiceArray}$
\[
        A = \begin{bNiceArray}{ccc|ccc}[columns-width = auto,first-row,last-col,code-for-first-row = \color{blue}\scriptstyle\rotate,code-for-last-col = \color{blue}\scriptstyle,nullify-dots,xdots/line-style=loosely dotted]
        1 & \Cdots & \lambda_1 & \lambda_1 + 1 & \Cdots & \lambda_1 + \lambda_2 \\
        0 & 1 & 0 & 0 & 0 & 0 & 1\\
        0 & \Ddots & 1 & 0 & 0 & 0 & \Vdots \\
        0 & 0 & 0 & 1 & 0 & 0 & \lambda_1 \\
        \hline
         & & & s & 1 & 0 & \lambda_1 + 1 \\
         & & & 0 & \Ddots & 1 & \Vdots \\
         & & & 0 & 0 & s & \lambda_1 + \lambda_2 
        \end{bNiceArray} 
\]
\end{comment}
% Denote by $R_i$ the $i$th row of $J$ 
%Let $g = E_{\lambda_2}\cdots E_1$ be the product of elementary matrices $E_i$ which correspond to the elementary row operations 
%$$R_{\lambda_1 + i - 1} \leftarrow R_{\lambda_1 + i-1} + \frac 1s R_{\lambda_1 + i}$$ 
%for $i = 1,\dots,\lambda_2$. Then $gJg^{-1} = A$. For the next part we introduce the notation $A_{\lambda_1+\lambda_2}$ for this matrix $A$.

@@

@@

% If we apply the change of basis $g$ to the ideal $I=I_1 + I_2$ and set $s = 0$ (if possible) do we recover the ideal of $\OO^{\lambda_1 + \lambda_2}$? Let $A^g = g(A_1 \oplus A_2) g^{-1}$...
% \end{remark}

@@

% TODO: place this better? 
% \acom{The following is a rough note.}

\begin{proposition}
    The family $\OO^{\lambda_1,\lambda_2}_{0,\AA}$ defined by \Cref{above} has zero fibre $\OO^\lambda$. 
\end{proposition}

We will show that this is a corollary of the following result on ``rank varieties''.

\begin{theorem}(\cite{eisenbud1989rank})
Let $r$ be a decreasing concave non-negative integral function with $r(0) = N$ (the dimension of the ambient vector space). Such a function is called a rank function. Then, in the space of $N\times N$ matrices, 
% The relevant result from Eisenbud--Saltman seems to be that 
$$X_r = X_{r,0} = \{A:\rk A^i \le r(i) \}$$ 
is the flat limit of 
% $$X_{r,\lambda} = \{A : \rk (A-\lambda_i)^j\le r(i,j)\}$$
$$X_{r,z} = \{A : \rk (A-z_i)^j\le r(i,j)\}$$
with $r(i,j)$ determined by $r(i)$.  
\end{theorem}

The case that we are interested in is the case of two eigenvalues: $(z_1,z_2) = (0,s)$. \acom{Does order matter?} 
% Let me change notation so that $\lambda_i$ are shapes and 
We begin by considering the open loci where the rank conditions are strict.

Recall that $\lambda_1 + \lambda_2 = \lambda$ is a partition of $N$. 
Let $m_i = \lambda_{i,1}$ ($i = 1,2$), $m = m_1 + m_2$, and define
$$r(j) = N - \#\text{~boxes in first $j$ columns of }\lambda$$
on $j = 1,\dots,m$. Note that $r(j-1) - r(j)$ is the length of the $j$th column of $\lambda$. As in \cite{eisenbud1989rank} we will denote this quantity $a_j$. 
%

% Again too small. The problem is I don't think that it is in general true that some permutation of the list $r(i-1) - r(i)$ will be the list of lengths of columns of $\lambda_1$ followed by the list of lengths of columns of $\lambda_2$ which is what we are looking for. For one thing there's in general more than one way to write a partition as a sum of two partitions. Ok, concretely, a third example. Take $\lambda_1 = (a,b)$ and $\lambda_2 = (c,d)$ then $\lambda = (a + c, b + d)$ and 
% $$r\{0,1,\dots,a + c\} = \{N, N-2, \dots, N-2(b+d), N-2(b+d) - 1,\dots, 0\}$$
% so 
% $$\{a_1,a_2,\dots\} = \{2,2,\dots,2,1,1,\dots,1\}$$
% with $b+d$ 2s and $a+c - b - d$ 1s. Wait a minute, nevermind, of course this is true. We have two partitions $\lambda_1$ and $\lambda_2$ and one way to form the partition $\lambda = \lambda_1 + \lambda_2$ is to break $\lambda_i$ down into columns, order the columns by length, and recombine. Cool. I think it's fine. 
% \acom{That's it for the note.}

@@

    Now assume that $ (L, s) \in \Gr^{\lambda_1, \lambda_2}_{0,s}$. We know that \acom{Rather than we know, it is by definition?} $ L(0) \in \Gr^{\lambda_1}$ and $L(s) \in \Gr^{\lambda_2} $.  This means that $t $ acting on 
    % $ \CC\xt^m / L(0)$ 
    $\cO^m/L(0)$ has Jordan type $ \lambda_1$ and $ t$ acting on 
    % $ \CC\xt[t-s]^m / L(s) $ 
    $\cO_s^m/L(s)$ has Jordan type $ \lambda_2$.  Now, the map 
    % $ L_0 / L \rightarrow \CC\xt^m / L(0)$ 
    $\CC[t]^m/L\to \cO^m/L(0)$ induces an isomorphism between the $0$-generalized eigenspace of $ t$ and $ \cO^m / L(0)$.   The same thing holds for the $s$-generalized eigenspace of $t $ and $ \cO_s^m/L(s)$. This shows that (i) implies (ii) and the logic can be reversed to see that (ii) implies (i). \acom{is this map $v + L \mapsto v + L(0)$?} \marginpar[]{*}
    % todo

@@

    the defining fact that lattices satisfying Equation~\ref{eq:defGrBDlambda} equivalently satisfy 
    \begin{equation}
        \begin{split}
            t\big|_{\CC\xt^m/L_1}\text{ has Jordan type} \le \lambda_1 \\
            t\big|_{\CC\xt^m/L_2}\text{ has Jordan type} \le \lambda_2 
        \end{split}
    \end{equation}
    where recall 
    $L_i = L\otimes \CC\xt$ ??? 
    % $L_i = L\otimes \CC[(t-p_i)^{-1}]$ 
    and $p_1 = s$ while $p_2 = 0$.

    $$\CC\xt^m/(L \otimes_{\CC[t]} \CC\xt) = \text{generalized $0$ eigenspace of $t$ on } \CC[t]^m/L$$
    carrying the action of $t $ to the action of $t$.

    The left hand side is the same thing as
    $$ \CC\xt^m / (L \otimes_{\CC[t]} \CC\xt) = (\CC[t]^m/L) \otimes_{\CC[t]} \CC\xt $$
    \acom{Somehow, restricting to an eigenspace is like inverting/forgetting the action of $t$ by any other generalized eigenvalue? Basic linear algebra? Joel?}
\end{enumerate}
    \end{enumerate}
\end{proof}

@@

    Need: operations of taking quotients and taking generalized eigenspaces (i.e.\ $-\otimes\CC\xt[t-a]$) commute.
    See $(*)$ above. Also 
    % 
    $$\CC\xt^m/(L \otimes_{\CC[t]} \CC\xt) = \text{generalized $0$ eigenspace of $t$ on } \CC[t]^m/L$$
    carrying the action of $t $ to the action of $t$.

    The left hand side is the same thing as
    $$ \CC\xt^m / (L \otimes_{\CC[t]} \CC\xt) = (\CC[t]^m/L) \otimes_{\CC[t]} \CC\xt $$
    \acom{Restricting to an eigenspace is like inverting (forgetting) the action of $t$ by any other generalized eigenvalue. Does this have to do with the fact that a functor $F: \sch \to \set$ commutes with projective limit if and only if it's representable?}

    Exercise 2.20 from \cite{atiyah2018introduction}: If $f : A \to B$ is a ring homomorphism and $M$ is a flat $A$-module, then $M_B = B\otimes_A M$ is a flat $B$-module. Take $A=\CC[t]$ and $B=\CC\xt$. \marginpar[]{*}
\end{proof}
